Reduced GW invariants from cuspidal curves

### 2018-01-05 draft_merged.tex
reviewing intro and sec 1 after Francesca's merge

\newcommand{\M}[4]{\overline{\pazocal M}_{#1,#2}(#3,#4)}
\newcommand{\Mone}[3]{\overline{\pazocal M}^{(1)}_{#1}(#2,#3)}
\newcommand{\PP}{\mathbb P}
\newcommand{\OO}{\mathcal O}
\renewcommand{\to}{\rightarrow}
\newcommand{\pr}{\rm{pr}}
\newcommand{\Aaff}{\mathbb A}
\newcommand{\A}{\mathcal A}
\newcommand{\B}{\mathcal B}
\newcommand{\X}{\pazocal X}
\newcommand{\XP}{\pazocal{X\!P}}
\newcommand{\tXP}{\widetilde{\pazocal{X\!P}}}
\renewcommand{\L}{\mathcal L}
\newcommand{\hL}{\hat{\mathcal L}}
\newcommand{\C}{\mathfrak C}
\newcommand{\cC}{\mathcal C}
\newcommand{\Z}{\pazocal Z}
\newcommand{\Zp}{\pazocal Z^p}
\newcommand{\tZ}{\widetilde{\pazocal Z}}
\newcommand{\tZp}{\widetilde{\pazocal Z}^p}
\newcommand{\MM}{\mathfrak M}
\newcommand{\pP}{\mathfrak P}
\newcommand{\hP}{\hat{\mathfrak P}}
\newcommand{\hM}{\hat{\mathfrak M}}
\newcommand{\oM}{\overline{\pazocal M}}
\newcommand{\tM}{\widetilde{\pazocal M}}
\newcommand{\R}{\operatorname{R}}
\newcommand{\Gm}{\mathbb{G}_{\rm{m}}}
\newcommand{\w}{\mathbf{w}}
\newcommand{\PtoM}{\rho}
\newcommand{\vir}[1]{[#1]^{\rm{vir}}}
\newcommand{\virloc}[1]{[#1]^{\rm{vir}}_{\rm{loc}}}

@@

\begin{abstract}
abstract
\end{abstract}

@@

The moduli stack of stable maps to projective space is a fundamental object in Gromov-Witten theory, since any moduli space of maps to a projective variety is cut inside the former by a number of equations induced by those of $X\subseteq \PP^r$. Yet its structure in higher genus is still largely mysterious.

Recall that in genus zero $\M{0}{n}{\PP^r}{d}$ is a smooth stack of the expected dimension. Furthermore for any split vector bundle $E=\oplus_{i} \mathcal O_{\PP^r}(l_i)$ the sheaf $\pi_*f^*E$ is a vector bundle on $\M{0}{n}{\PP^r}{d}$.
\bcd
\mathcal C_{0,n}(\PP^r,d)\ar[r,"f"]\ar[d,"\pi"] & \PP^r\times \M{0}{n}{\PP^r}{d} \\
\M{0}{n}{\PP^r}{d} &
\ecd

It is a matter of functoriality of virtual fundamental classes \cite{KON} \cite{CKL} \cite{KKP} that, if $i\colon X\hookrightarrow \PP^r$ is a complete intersection of degree $(l_1,\ldots,l_k)$ and we denote by $j$ the corresponding inclusion of moduli spaces of maps,
%$$\bigsqcup_{\beta\in A_1(X)\colon i_*(\beta)=d}\M{0}{n}{X}{\beta}\hookrightarrow\M{0}{n}{\PP^r}{d}$$
then the following equality holds:%, which is expected from the interpretation of Chern classes as zero loci of sections (if $X$ is cut by a generic section $s\in H^0(\PP^r,E)$, then the induced section $\tilde s=\pi_*f^*(s)$ is also generic):
$$j_*[\bigsqcup_{\beta\in A_1(X)\colon i_*(\beta)=d}\M{0}{n}{X}{\beta}]^{\operatorname{vir}}=c_{\operatorname{top}}(E)\cap[\M{0}{n}{\PP^r}{d}].$$
This so-called ``hyperplane property'' allows us to compute GW invariants of complete intersections as twisted invariants of projective space, which enables us to perform computations by localisation.

The situation in higher genera is much more intricate: the moduli space $\M{g}{n}{\PP^r}{d}$ is neither irreducible, nor pure dimensional, and the sheaf $\pi_*f^*\mathcal O_{\PP^r}(l)$ is not a vector bundle. The enumerative meaning of the invariants is affected by the appearance of unexpected components, representing degenerate contributions from curves of lower genera.

The situation in genus $1$ is better understood: the geometry of the moduli space of genus one stable maps to projective space $\M{1}{n}{\PP^r}{d}$ has been widely studied since \cite{Vre}. Contrary to higher genus, the closure of the locus of maps the source of which is a smooth elliptic curve is an irreducible component of $\M{1}{n}{\PP^r}{d}$, which is usually called the \emph{main component}.
%One of the reasons for this (in fact the only reason in genus one) is that there are maps from reducible curves with a genus $g$ core that is contracted and some rational tails that carry all the degree; if $f\colon C\to \PP^r$ is such a map, then $H^1(C,f^*\mathcal O_{\PP^r}(1))\neq 0$, as can be easily seen from the normalisation sequence.

In an attempt to extract the enumerative information hidden within Kontsevich's moduli space, Vakil and Zinger managed to desingularise the main component by performing a sequence of blow-ups on an underlying space of weighted curves \cite{VZpreview} \cite{VZ} (see also \cite{HL}). The end result $\widetilde{\pazocal M}_{1,n}(\PP^r,d)^{\rm{main}}$ is a stack on which $\tilde\pi_*\tilde f^*\mathcal O_{\PP^r}(l)$ is a vector bundle, which allows them to define a notion of \emph{reduced invariants} for complete intersections in $\PP^r$, based on the hyperplane property:  %The idea is that blowing up successively the locus of degree zero elliptic curve with one rational tail (\emph{elliptic tails}), then with two (\emph{elliptic bridges}), etc. and taking the fiber product
%\bcd
%\widetilde{\mathcal M}_{1,n}(\PP^r,d)\ar[r]\ar[d] & \M{1}{n}{\PP^r}{d}\ar[d] \\
%\widetilde {\mathfrak M}_{1,n}^{\operatorname{wt}=d} \ar[r] & \mathfrak M_{1,n}^{\operatorname{wt}=d}
%\ecd
%the strict transform of the main component becomes smooth and meets the rest of the components in normal crossing fashion (the \'{e}tale local model being the union of two linear subspaces of different dimensions in affine space). Furthermore, the sheaves $\tilde\pi_*\tilde f^*\mathcal O_{\PP^r}(l)$ contain a vector subbundle of rank $dl$ supported on the strict transform of the main component. This paved the way to the definition of reduced invariants.
\begin{dfn}
Let $X$ be a complete intersection of degree $(l_1,\ldots,l_k)$ in $\PP^r$. The genus one reduced Gromov-Witten invariants of $X$ are defined by integrating against: $$[\widetilde{\pazocal M}_{1,n}(\PP^r,d)^{\rm{main}}]\cap c_{\rm{top}}(\tilde\pi_*\tilde f^*(\oplus_{i=1}^k\mathcal O_{\PP^r}(l_i))).$$
\end{dfn}

Recently there has been a different attempt to discard some of the unwanted components. Viscardi \cite{VISC} has introduced a series of \emph{alternate compactifications} of the moduli space of maps from elliptic curves, based on Smyth's work on the birational geometry of $\oM_{1,n}$. In Viscardi's space $\oM_{1,n}^{(m)}(X,\beta)$ maps that contract an elliptic curve with at most $m$ rational tails are destabilised, and replaced by maps from $l$-elliptic fold singularities with $l\leq m$.

The main result of this paper is that reduced invariants of the quintic three-fold can be recovered from Viscardi's alternate compactification.

\begin{thm}
Let $X$ be a generic smooth quintic three-fold. Then:
\[
N_1^{\rm{red}}(X,d)=N_1^{\rm{cusp}}(X,d)
\]
where:
\[N_1^{\rm{red}}(X,d):=\deg \left([\widetilde{\pazocal M}_{1}(\PP^4,d)^{\rm{main}}]\cap c_{\rm{top}}(\tilde\pi_*\tilde f^*(\mathcal O_{\PP^4}(5)))\right)\]
and
\[N_1^{\rm{cusp}}(X,d):=\deg [\Mone{1}{X}{d}]^{\rm{vir}}.
\]
\end{thm}

The insight for this fact comes from the following formula, the proof of which was first given by J. Li and A. Zinger \cite{LZ} in the symplectic category:

$$N_1(X,d)=N_1^{\rm{red}}(X,d)+\frac{1}{12}N_0(X,d).$$
Noticing that the genus $0$ contribution to this formula only comes from the boundary component with a single rational tail of degree $d$ suggests that removing such component would provide a more direct approach to reduced invariants. This is precisely what Viscardi's space of 1-stable maps does.

\textcolor{red}{Li-Zinger's formula was later proved in the algebro-geometric setting by H.-W. Chang and J. Li in \cite{CL}. This was only the final of a series of papers where they developed the theory of stable maps with $p$-fields and applied it together with the techniques of cosection-localised virtual classes \cite{KLcosection}. $p$-fields provide a way around the fact that we cannot control the geometry of $\oM_1(X,d)$ and there is no compatible perfect obstruction theory relative to the inclusion $\oM_1(X,d)\hookrightarrow \oM_1(\PP^4,d)$; indeed Chang and Li managed to include the unwanted part of the obstruction theory in the moduli space, and by these means endow a well-understood space (very similar to $\oM_1(\PP^4,d)$) with an obstruction theory counting genus $1$ curves in the quintic.}
\textcolor{blue}{To prove Li-Zinger formula in the algebro-geometric setting, one would like to understand how the virtual class of  $\oM_1(X,d)$ splits on the various component of the moduli space, but this seems to be unaccesible. To overcome this difficulty, Chang Li developed the theory of stable maps with $p$- fields.}

The proof of our main theorem follows from adapting the techniques of Chang and Li to Viscardi's moduli spaces. We show the following:
\begin{prop}
There exists a well-defined $1$-stabilisation morphism at the level of weighted pre-stable curves:
\[
\MM_{1,n}^{\rm{wt}}\to \MM_{1,n}^{(1),\rm{wt}}
\]
\end{prop}
The fiber product
\[
\MM_{1,n}^{\rm{wt}}\times_{\MM_{1,n}^{(1),\rm{wt}}}\Mone{1}{\PP^4}{d}^p
\]
is endowed with a virtual class counting $1$-stable maps to the quintic three-fold (the arguments in \cite{CLpfields} carry over almost unchanged). On the other hand we find a way to compare the main contribution of this space with Vakil-Zinger reduced invariants. A dimensional computation shows that this is the only enumeratively meaningful component.

\subsection*{Outline of the paper}
maybe table of notations
\section{Genus one stable maps to $\PP^r$ - Equations, components and alternate compactifications}

\subsection{Local equations and components}
We start by recalling a description of the global geometry of $\M{1}{n}{\PP^r}{d}$. Besides the main component, which we have defined in the Introduction, for every positive integer $k$ and partitions $\lambda\vdash d$ into $k$ parts and $\mu\vdash n$ into $k$ parts such that no $(\lambda_i,\mu_i)_{i=1}^k$ is either $(0,1)$ or $(0,2)$, there is a \emph{boundary} irreducible component $D_{\lambda,\mu}(\PP^r,d)$ defined to be the closure of the locus where:
\begin{description}
\item[i] the source curve is obtained by gluing a smooth elliptic curve with $k$ many $\PP^1$'s as rational tails,
\item[ii] the map contracts the core elliptic curve to a point, and
\item[iii] the rational tail $R_i$ has degree $\lambda_i$ and $\mu_i$-many marked points.
\end{description}
Indeed $D_{\lambda,\mu}(\PP^r,d)$ is the image of the gluing morphism from the fiber product:
$$\M{1}{k}{\PP^r}{0}\times_{\PP^r}\prod_{i=1}^k\M{0}{\mu_i+1}{\PP^r}{\lambda_i}.$$

The minimal arithmetic genus $1$ subcurve is called the \emph{core}. We then have:
\begin{prop}\label{prop:components}
\emph{(1)} The ones that we have described above are all the irreducible components of $\M{1}{n}{\PP^r}{d}$, %The set of irreducible components of $\M{1}{n}{\PP^r}{d}$ is in bijection with $\{\text{stable}\ (\lambda\vdash d,\mu\vdash n)|\ \#(\text{parts})(\lambda)=\#(\text{parts})(\mu)\}\cup\{\rm{main}\}$,
i.e.
$$\M{1}{n}{\PP^r}{d}=\M{1}{n}{\PP^r}{d}^{\rm{main}}\cup\bigcup_{\lambda,\mu}D_{\lambda,\mu}(\PP^r,d).$$

\emph{(2)} A map $[f]$ lies in the boundary of the main component  if and only if:
\begin{itemize}
\item either $f$ is non-constant on an irreducible component of the core,
\item or if $(E\ {}_{\mathbf p}\!\sqcup_{\mathbf q}\bigsqcup_{i=1}^k R_i,f)$ is a stable map with $k$ rational tails, where $E$ is the maximal contracted genus one subcurve, then $\{\operatorname{d}\!f(T_{q_i}R_i)\}_{i=1}^k$ are \emph{linearly dependent} in $T_{f(E)}\PP^r$.
\end{itemize}
\end{prop}

@@

%In order to understand how the Gromov-Witten invariants differ from the naive expectation of counting smooth embedded genus $1$ curves in the given homology class, one needs a good understanding of the virtual fundamental class and of its splitting on the various components. 
%This has led to the study of explicit local equation for the moduli space of stable maps in a smooth (relative to a moduli space of decorated curves) ambient space \cite{HL}, which we shall now briefly explain.
We now review Hu-Li's procedure \cite{HL} to find local equations of $\M{1}{n}{\PP^r}{d}$ in a smooth ambient space. These are useful in describing the structure of the intrinsic normal cone and its splitting.

Recall that a map $C\to\PP^r$ is given by a line bundle $L$ on $C$ together with $r+1$ sections in $H^0(C,L)$ that generate the line bundle. It is therefore natural to embed the space of stable maps as an open inside $\pi_*\mathcal L^{\oplus r+1}$ on the universal Picard stack $\mathfrak{P}_{1,n}=\mathfrak{Pic}(\cC_{1,n}\to\MM_{1,n})$.
Hu and Li actually work on a local chart of the algebraic stack $\mathfrak M_{1,n}^{\rm{div}}$, parametrising families of pre-stable curves with a relative Cartier divisor of degree $d$. 
 %The problem is that this cone is not smooth, due to the jumps in the dimension of the fiber at the boundary of the moduli space.
%(besides there is a small issue due to the $\mathbb G_m$ automorphism group of objects of $\mathfrak{P}_{1,n}$, which can be avoided by dishomogenising).

Let $[f\colon C\to \PP^r]$ be a point of $\M{1}{n}{\PP^r}{d}$: we may fix homogeneous coordinates on $\PP^r$ in such a way that $D_0=f^{-1}\{x_0=0\}$ is a simple smooth divisor on $C$ (i.e. a $d$-uple of smooth points); this property will then hold in a neighbourhood of $[f]$. This gives a map from an \'{e}tale chart $\pazocal U$ of $\M{1}{n}{\PP^r}{d}$ to an \'{e}tale chart $\pazocal V$ of $\mathfrak M_{1,n}^{\rm{div}}$, that we may assume to land in the locus where the divisor is made of $d$ distinct smooth points. Notice that this locus is smooth over $\mathfrak M_{1,n}$ by the deformation theory of smooth subschemes. A morphism to $\PP^r$ shall now be thought of as a curve-divisor pair $(C,D)$ together with $r$ sections of $\mathcal O_C(D)$: the morphism can be written as $[1:u_1:\ldots:u_r]$, where $1$ means the given section of $\mathcal O_C\to\mathcal O_C(D)$.

>>>>

\newcommand{\M}[4]{\overline{\pazocal M}_{#1,#2}(#3,#4)}
\newcommand{\Mone}[3]{\overline{\pazocal M}^{(1)}_{#1}(#2,#3)}
\newcommand{\PP}{\mathbb P}
\renewcommand{\k}{\mathbf k}
\newcommand{\OO}{\mathcal O}
\renewcommand{\to}{\rightarrow}
\newcommand{\pr}{\rm{pr}}
\newcommand{\Aaff}{\mathbb A}
\newcommand{\A}{\pazocal A}
\newcommand{\B}{\pazocal B}
\newcommand{\X}{\pazocal X}
\newcommand{\XP}{\pazocal{X\!P}}
\newcommand{\tXP}{\widetilde{\pazocal{X\!P}}}
\renewcommand{\L}{\mathcal L}
\newcommand{\hL}{\hat{\mathcal L}}
\newcommand{\C}{\mathfrak C}
\newcommand{\cC}{\mathcal C}
\newcommand{\Z}{\pazocal Z}
\newcommand{\Zp}{\pazocal Z^p}
\newcommand{\tZ}{\widetilde{\pazocal Z}}
\newcommand{\tZp}{\widetilde{\pazocal Z}^p}
\newcommand{\MM}{\mathfrak M}
\newcommand{\pP}{\mathfrak P}
\newcommand{\hP}{\hat{\mathfrak P}}
\newcommand{\hM}{\hat{\mathfrak M}}
\newcommand{\oM}{\overline{\pazocal M}}
\newcommand{\tM}{\widetilde{\pazocal M}}
\newcommand{\R}{\operatorname{R}}
\newcommand{\Gm}{\mathbb{G}_{\rm{m}}}
\newcommand{\w}{\mathbf{w}}
\newcommand{\PtoM}{\rho}
\newcommand{\vir}[1]{[#1]^{\rm{vir}}}
\newcommand{\virloc}[1]{[#1]^{\rm{vir}}_{\rm{loc}}}

@@

\begin{abstract}

We prove that the $1$-stabilisation morphism, replacing elliptic tails
with cusps, is well-defined at the level of genus $1$ weighted-stable
curves. We use this to endow a substack of $\oM_1(\PP^4,d)$ with a
virtual class that ignores the boundary contribution of maps with a
contracted elliptic tail. We thus prove that Vakil-Zinger's reduced
invariants of the quintic threefold can be computed as Smyth-Viscardi's
$1$-stable (i.e. cuspidal) invariants. To do so we adapt Jun Li's
techniques of $p$-fields and local equations for the moduli space.

\end{abstract}

@@

The moduli stack of stable maps to the projective space $\PP^r$ is of fundamental importance in Gromov-Witten theory, since any moduli space of maps to a projective variety $X\subseteq \PP^r$ is cut inside the former by a number of equations induced by generators of $\pazocal I_{X/\PP^r}$. Yet its structure in higher genus is still largely mysterious.

Recall that in genus zero $\M{0}{n}{\PP^r}{d}$ is a smooth stack of the expected dimension $r-3+d(r+1)+n$. Furthermore the rational Gromov-Witten theory of a complete intersection $i\colon X\hookrightarrow \PP^r$ of degree $(l_1,\ldots,l_k)$ can be determined as a twisted theory for $\PP^r$ and thus computed via localisation: for any split vector bundle $E=\oplus_{i} \mathcal O_{\PP^r}(l_i)$ the sheaf $\pi_*f^*E$ is a vector bundle on $\M{0}{n}{\PP^r}{d}$, where we have denoted by $\pi$ the projection from the universal curve and by $f$ the universal stable map.
%\bcd
%\mathcal C_{0,n}(\PP^r,d)\ar[r,"f"]\ar[d,"\pi"] & \PP^r%\times \M{0}{n}{\PP^r}{d} \\
%\M{0}{n}{\PP^r}{d} &
%\ecd
It is a matter of functoriality of virtual fundamental classes \cite{KON} \cite{CKL} \cite{KKP} that the following \emph{hyperplane property} holds
%$$\bigsqcup_{\beta\in A_1(X)\colon i_*(\beta)=d}\M{0}{n}{X}{\beta}\hookrightarrow\M{0}{n}{\PP^r}{d}$$
%then the following equality holds:, which is expected from the interpretation of Chern classes as zero loci of sections (if $X$ is cut by a generic section $s\in H^0(\PP^r,E)$, then the induced section $\tilde s=\pi_*f^*(s)$ is also generic):
$$j_*[\bigsqcup_{\beta\in A_1(X)\colon i_*(\beta)=d}\M{0}{n}{X}{\beta}]^{\operatorname{vir}}=c_{\operatorname{top}}(E)\cap[\M{0}{n}{\PP^r}{d}].$$
%This, so-called ``hyperplane property'', allows us to compute GW invariants of complete intersections as twisted invariants of projective space, where computations can be done via localisation.

For $g\geq 1$ the moduli space $\M{g}{n}{\PP^r}{d}$ is neither irreducible, nor pure dimensional, and the sheaf $\pi_*f^*\mathcal O_{\PP^r}(l)$ is not a vector bundle. The enumerative meaning of the invariants is spoilt by degenerate contributions from curves of lower genera, represented by boundary components of the moduli space where a subcurve of positive genus is contracted.

The situation in genus $1$ is better understood: the geometry of $\M{1}{n}{\PP^r}{d}$ has been widely studied since \cite{Vre}. Contrary to higher genus, the closure of the locus of maps the source of which is a smooth elliptic curve is an irreducible component of $\M{1}{n}{\PP^r}{d}$, which is usually called the \emph{main component}.
Its boundary points were characterised in \cite{Vre} as follows: $[f\colon C\to \PP^r]\in\oM_1(\PP^r,d)$ is smoothable, i.e. it can be made into the special fiber of a $1$-dimensional family the generic fiber of which is a non-constant map from a smooth elliptic curve, if and only if either $f$ is not constant on the minimal genus $1$ subcurve $E$ of $C$, or the image of the tangent vectors to the rational tails are linearly dependent in $T_{f(E)}\PP^r.$  In Section~\ref{sec:section1} we give a new proof of this fact~\ref{prop:smoothability} using Hu-Li's local equations for the moduli space; furthermore we provide a different characterisation of the smoothable maps as those that factor through a non-degenerate map from a Cohen-Macaulay genus~$1$ singularity \ref{prop:smoothability2}, which we classify extending an argument due to Smyth.
%One of the reasons for this (in fact the only reason in genus one) is that there are maps from reducible curves with a genus $g$ core that is contracted and some rational tails that carry all the degree; if $f\colon C\to \PP^r$ is such a map, then $H^1(C,f^*\mathcal O_{\PP^r}(1))\neq 0$, as can be easily seen from the normalisation sequence.

From the enumerative point of view, it would be ideal to isolate the contribution to the invariants given by the main component.
In an attempt to extract this information hidden within Kontsevich's moduli space, 
Vakil and Zinger managed to desingularise the main component by performing a sequence of blow-ups on an underlying space of weighted curves \cite{VZpreview} \cite{VZ} (see also \cite{HL}). The end result $\widetilde{\pazocal M}_{1,n}(\PP^r,d)^{\rm{main}}$ is a smooth and irreducible stack on which $\tilde\pi_*\tilde f^*\mathcal O_{\PP^r}(l)$ is a vector bundle. This fact allowed them to define a notion of \emph{reduced invariants} for complete intersections in $\PP^r$ by mimicking the hyperplane property of the genus $0$ theory. %The idea is that blowing up successively the locus of degree zero elliptic curve with one rational tail (\emph{elliptic tails}), then with two (\emph{elliptic bridges}), etc. and taking the fiber product
%\bcd
%\widetilde{\mathcal M}_{1,n}(\PP^r,d)\ar[r]\ar[d] & \M{1}{n}{\PP^r}{d}\ar[d] \\
%\widetilde {\mathfrak M}_{1,n}^{\operatorname{wt}=d} \ar[r] & \mathfrak M_{1,n}^{\operatorname{wt}=d}
%\ecd
%the strict transform of the main component becomes smooth and meets the rest of the components in normal crossing fashion (the \'{e}tale local model being the union of two linear subspaces of different dimensions in affine space). Furthermore, the sheaves $\tilde\pi_*\tilde f^*\mathcal O_{\PP^r}(l)$ contain a vector subbundle of rank $dl$ supported on the strict transform of the main component. This paved the way to the definition of reduced invariants.

\begin{dfn}
Let $X$ be a complete intersection of degree $(l_1,\ldots,l_k)$ in $\PP^r$. The genus $1$ reduced invariants of $X$ are defined by integrating against: $$ c_{\rm{top}}(\tilde\pi_*\tilde f^*(\oplus_{i=1}^k\mathcal O_{\PP^r}(l_i)))\cap[\widetilde{\pazocal M}_{1,n}(\PP^r,d)^{\rm{main}}].$$
\end{dfn}

Recently there has been a different attempt to discard some of the boundary components. Based on Smyth's work on the birational geometry of $\oM_{1,n}$ \cite{SMY1,SMY2}, Viscardi \cite{VISC} has introduced a series of \emph{alternate compactifications} of the moduli space of maps from smooth elliptic curves. Compared to Kontsevich's space of stable maps, in Viscardi's space $\oM_{1,n}^{(m)}(X,\beta)$ maps that contract a genus $1$ curve with at most $m$ rational tails are destabilised, and replaced by maps from $l$-elliptic singularities with $l\leq m$.

The main result of the present paper is that reduced invariants of the quintic threefold can be recovered from Viscardi's alternate compactifications.

\begin{thm}
Let $X_5\subseteq\PP^4$ be a generic smooth quintic threefold. Then:
\[
N_1^{\rm{red}}(X_5,d)=N_1^{\rm{cusp}}(X_5,d),
\]
where:
\[N_1^{\rm{red}}(X_5,d):=\deg \left(c_{\rm{top}}(\tilde\pi_*\tilde f^*(\mathcal O_{\PP^4}(5)))\cap [\widetilde{\pazocal M}_{1}(\PP^4,d)^{\rm{main}}]\right)\]
and
\[N_1^{\rm{cusp}}(X_5,d):=\deg [\Mone{1}{X_5}{d}]^{\rm{vir}}.
\]
\end{thm}

Below we will suppress the subscript $5$ from $X_5$. The insight for this theorem comes from the following formula, the proof of which was first given by J. Li and A. Zinger \cite{LZ} in the symplectic category:
\begin{equation}\label{eq:VakilZinger}
N_1(X,d)=N_1^{\rm{red}}(X,d)+\frac{1}{12}N_0(X,d).
\end{equation}
The first proof of Li-Zinger's formula in the algebro-geometric setting is due to H.-W. Chang and J. Li in \cite{CL}. Their splitting of the intrinsic normal cone makes it patent that the genus $0$ contribution to the right hand side only comes from the boundary component with a single rational tail of degree $d$. This suggests that removing such a component would provide a more direct approach to reduced invariants for the quintic threefold. This is precisely what Viscardi's space of $1$-stable maps does.

%To prove the result we essentially need to understand how the various component of the moduli space contribute to the invariants, so we have to look at the algebraic techniques that  have been used to prove~\eqref{eq:VakilZinger}.

%Indeed, after the symplectic proof, Li-Zinger's formula was proved . The latter is only the final a series of papers where they developed the theory of stable maps with $p$-fields and applied it together with the techniques of cosection-localised virtual classes \cite{KLcosection}. $p$-fields provide a way around the fact that we cannot control the geometry of $\oM_1(X,d)$ and there is no compatible perfect obstruction theory relative to the inclusion $\oM_1(X,d)\hookrightarrow \oM_1(\PP^4,d)$; indeed Chang and Li managed to include the unwanted part of the obstruction theory in the moduli space, and by these means endow a well-understood space (very similar to $\oM_1(\PP^4,d)$) with an obstruction theory counting genus $1$ curves in the quintic.
The proof of our main theorem follows from adapting the techniques of Chang and Li to Viscardi's moduli spaces. Here is an outline of the proof.

%Alternatively, if one does not want to use the theory of stable maps with fields, the idea would be to follow the alternative proof of~\eqref{eq:VakilZinger} recently proposed by \cite{preprint}. In that case, they still consider the Hu-Li blow up of the moduli space of weighted curve and they then look at $\tM_1(X)\hookrightarrow \tM_1(\PP^4) $, where $X$ is the quintic threefold and study directly the cone of $\tM_1(X).$ Following~\cite{preprint} steps, we can attack directly $\tM_1(X)^{(1)}$ and show that only the main component has a non trivial contribution to the invariants.

%In both cases the comparison with the genus~1 invariants comes from the following comparison result.
%We show the following:
We first consider a moduli space of $1$-stable maps with $p$-fields $\Mone{1}{\PP^4}{d}^p$. In the work of Chang and Li, $p$-fields provide a way around the fact that we cannot control the geometry of $\oM_1(X,d)$ and there is no compatible perfect obstruction theory relative to the inclusion $\oM_1(X,d)\hookrightarrow \oM_1(\PP^4,d)$. By introducing the moduli space of maps with $p$-fields the authors manage to include the unwanted part of the obstruction theory into the moduli space, and to endow a well-understood space (indeed very similar to $\oM_1(\PP^4,d)$) with an obstruction theory counting genus $1$ curves in the quintic. Paralleling their proof, i.e. using cosection localised virtual cycles \cite{KLcosection} and the deformation to the normal cone $(X_5\subseteq \PP^4)\rightsquigarrow(X_5\subseteq N_{X/\PP^4})$, we see that:
\begin{prop}
$\deg\virloc{\Mone{1}{\PP^4}{d}^p}=\deg\vir{\Mone{1}{X}{d}}.$
\end{prop}
Indeed it is enough to notice that all the constructions in \cite{CLpfields} carry over to the setting of a family of Gorenstein (not necessarily nodal) curves.

We then need the following comparison morphism.
\begin{prop}
There exists a well-defined $1$-stabilisation morphism at the level of weighted-stable curves:
\[
\MM_{1,n}^{\rm{wt,st}}\to \MM_{1,n}^{\rm{wt,st}}(1)
\]
\end{prop}
The fiber product
\[
\Z:=\MM_{1,n}^{\rm{wt,st}}\times_{\MM_{1,n}^{\rm{wt,st}}}\Mone{1}{\PP^4}{d}
\]
is a substack of $\oM_1(\PP^4,d)$ avoiding the boundary component with one rational tail of degree $d$.

It is harder to compare the fiber product
\[
\Zp:=\MM_{1,n}^{\rm{wt,st}}\times_{\MM_{1,n}^{\rm{wt,st}}}\Mone{1}{\PP^4}{d}
\]
with $\oM_1(\PP^4,d)^p$. Yet $\Zp$ is endowed with a cosection-localised virtual class counting $1$-stable maps to the quintic threefold (by cosection-localised virtual pullback). The same splitting arguments as in \cite{CLpfields} allow us to evince a main contribution which coincides with Vakil-Zinger's reduced invariants. On the other hand a dimensional computation shows that this is the only enumeratively meaningful component.

\subsection*{Future directions}

\subsection*{Outline of the paper}

\subsection*{Table of notations} We work over an algebraically closed field $\k$ of characteristic $0$.

$T$ is the spectrum of a discrete valuation ring (DVR) with closed point $0$ and generic point $\eta$.

Subcurves are always connected.

\section{$\oM_1(\PP^r,d)$ - Components, equations  and alternate compactifications}\label{sec:section1}

\subsection{Local equations and components}
We start by recalling a description of the global geometry of $\oM_1(\PP^r,d)$. Besides the main component, which was defined in the Introduction, for every positive integer $k$ and partition $\lambda\vdash d$ into $k$ parts such that no $\lambda_i=0$, there is an irreducible \emph{boundary component} $D_{\lambda}(\PP^r,d)$ defined to be the closure of the locus where:
\begin{description}
\item[i] the source curve is obtained by gluing a smooth elliptic curve $E$ with $k$ many $\PP^1$'s as rational tails $R_i$,
\item[ii] the map contracts the elliptic curve $E$ to a point, and
\item[iii] the map has degree $\lambda_i$ on the rational tail $R_i$.
\end{description}
Indeed $D_{\lambda}(\PP^r,d)$ is the image of the gluing morphism from the fiber product:
$$\M{1}{k}{\PP^r}{0}\times_{\PP^r}\prod_{i=1}^k\M{0}{1}{\PP^r}{\lambda_i}.$$

\noindent The \emph{minimal} arithmetic genus $1$ subcurve is called the \emph{core} or the \emph{circuit}. %We then have:
\begin{prop}\label{prop:components}
\emph{(1)} All the irreducible components of $\oM_1(\PP^r,d)$ have been described above: %The set of irreducible components of $\M{1}{n}{\PP^r}{d}$ is in bijection with $\{\text{stable}\ (\lambda\vdash d,\mu\vdash n)|\ \#(\text{parts})(\lambda)=\#(\text{parts})(\mu)\}\cup\{\rm{main}\}$,
\[\oM_1(\PP^r,d)=\oM_1(\PP^r,d)^{\rm{main}}\cup\bigcup_{\lambda}D_{\lambda}(\PP^r,d).\]

\emph{(2)} A map $[f]$ lies in \emph{the boundary of the main component}  if and only if:
\begin{itemize}
\item either $f$ is non-constant on at least one irreducible component of the core,
\item or if $(E\ {}_{\mathbf p}\!\sqcup_{\mathbf q}\bigsqcup_{i=1}^k R_i,f)$ is a stable map with $k$ rational tails, where $E$ is the \emph{maximal} contracted genus $1$ subcurve, then $\{\operatorname{d}\!f(T_{q_i}R_i)\}_{i=1}^k$ are \emph{linearly dependent} in $T_{f(E)}\PP^r$.
\end{itemize}
\end{prop}

@@

%In order to understand how the Gromov-Witten invariants differ from the naive expectation of counting smooth embedded genus $1$ curves in the given homology class, one needs a good understanding of the virtual fundamental class and of its splitting on the various components. 
%This has led to the study of explicit local equation for the moduli space of stable maps in a smooth (relative to a moduli space of decorated curves) ambient space \cite{HL}, which we shall now briefly explain.
We now review Hu-Li's procedure for finding local equations of $\oM_1(\PP^r,d)$ in a smooth ambient space \cite{HL}. They will be useful when describing the structure of the intrinsic normal cone and its splitting.

Recall that a map $C\to\PP^r$ is given by a line bundle $L$ on $C$ together with $r+1$ sections in $H^0(C,L)$ that generate the line bundle. It is therefore natural to embed the space of stable maps as an open inside $\pi_*\mathcal L^{\oplus r+1}$ on the universal Picard stack $\mathfrak{P}_{1}=\mathfrak{Pic}(\cC_{1}\to\MM_{1})$.
Hu and Li actually work over a local chart of the algebraic stack $\mathfrak M_{1}^{\rm{div}}$, parametrising families of prestable curves with a relative Cartier divisor (of total degree $d$).
 %The problem is that this cone is not smooth, due to the jumps in the dimension of the fiber at the boundary of the moduli space.
%(besides there is a small issue due to the $\mathbb G_m$ automorphism group of objects of $\mathfrak{P}_{1,n}$, which can be avoided by dishomogenising).

Let $[f\colon C\to \PP^r]$ be a point of $\oM_1(\PP^r,d)$: we may fix homogeneous coordinates on $\PP^r$ in such a way that $D_0:=f^{-1}\{x_0=0\}$ is a simple divisor contained supported on the smooth locus of $C$ (i.e. a $d$-uple of smooth points); this property will then hold in a neighbourhood of $[f]$. This gives a morphism from an \'{e}tale chart $\pazocal U$ of $\oM_1(\PP^r,d)$ to an \'{e}tale chart $\pazocal V$ of $\mathfrak M_{1}^{\rm{div}}$, that we may assume to land in the locus where the divisor consists of $d$ distinct smooth points. Notice that this locus is smooth over $\mathfrak M_{1}$ by the deformation theory of smooth subschemes. A map to $\PP^r$ shall now be thought of as a curve-divisor pair $(C,D)$ together with $r$ sections of $\mathcal O_C(D)$: the map can be written as $[1:u_1:\ldots:u_r]$, where $1$ denotes the given section of $\mathcal O_C\to\mathcal O_C(D)$.
